---
title: "Question 2"
author: "Liz Igweonu"
date: "2023-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Introduction

```{r Data Exploration}
# Install packages 
# install.packages(c("ggplot2", "palmerpenguins", "janitor", "dplyr", "ragg", "ggpubr"))

# Load the packages
library(ggplot2)
library(palmerpenguins)
library(janitor)
library(dplyr)
library(ragg)
library(ggpubr)

# Load the functions definitions
source("functions/cleaning.r")

# Save the raw data:
write.csv(penguins_raw, "data/penguins_raw.csv")

# Check the raw data:
names(penguins_raw)

# Clean the data:
penguins_clean <- penguins_raw %>%
  clean_column_names() %>%
  shorten_species() %>%
  remove_empty_columns_rows()

# Check the cleaned data: 
names(penguins_clean)

# Save cleaned data:
write.csv(penguins_clean, "data/penguins_clean.csv")

# Filter the data for variables of interest
flipper_mass_data <- penguins_clean %>%
  subset_columns(c("flipper_length_mm", "body_mass_g", "species")) %>%
  remove_NA()

head(flipper_mass_data)

# Exploratory figure

results_scatterplot <- ggplot(flipper_mass_data, aes(x=flipper_length_mm,y=body_mass_g, color=species, shape = species)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(x = "Flipper Length (mm)",
         y = "Body Mass (g)",
       color = "Penguin Species",
       shape = "Penguin Species") +
  theme_light() +
  theme(legend.position = "bottom")

results_scatterplot

# Save the figure
agg_png("figures/exploratory.png",
        width = 30, height = 25,
        units = "cm",
        res = 600,
        scaling = 1.8)
results_scatterplot
dev.off()

```

### Hypothesis

Null: There is no correlation between penguin flipper length and body mass.

Alternative: There is a positive correlation between penguin flipper length and body mass.

### Statistical Methods

Pearson's correlation coefficient is used to test if there is a correlation between penguin flipper length and body mass. 

```{r Statistics}

# Pearson correlation coefficient 
test <- cor.test (flipper_mass_data$flipper_length_mm, flipper_mass_data$body_mass_g, method = "pearson")

test

# Coefficient of determination

r2 <- cor(flipper_mass_data$flipper_length_mm, flipper_mass_data$body_mass_g)^2

r2 <- r2*100

r2

# Pearson correlation coefficient grouped by species

speciestest <- flipper_mass_data %>%
  group_by(species) %>%
  summarise(cor=cor(flipper_length_mm, body_mass_g))

speciestest

```

### Results & Discussion

In the exploratory figure, with an increase in flipper length, the body mass is also increasing. There is a wide range for both flipper length and body mass and data is not concentrated at one point. Gentoos have both a higher flipper length and body mass compared to the other penguin species. Species Adelie and Chinstrap overlap in flipper length and body mass range.  

The result from the Pearson correlation coefficient test indicate there is a strong, positive correlation between flipper length and body mass as a value of R = 0.87 was obtained for all groups. This is highly significant since the p-value is < 2.2e-16. 75.9% of the variation in body mass is explained by flipper length (with an R^2 of 0.759). Therefore, we reject our null hypothesis and accept the alternative hypothesis.  

When grouped by species, Adelie has low correlation (0.47), Gentoo has high correlation (0.7) and Chinstrap has moderate correlation (0.64). This is highly significant for all groups and so there is strong evidence that there is a positive correlation between flipper length and body mass for each species. 

```{r Plotting Results}

# Results figure

results_cor <- ggscatter(flipper_mass_data, x = 'flipper_length_mm', y = 'body_mass_g',
          add = "reg.line", conf.int = TRUE,
          add.params = list(color = "blue",
                            fill = "lightgray"),
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Flipper Length (mm)", ylab = "Body Mass (g)")

results_cor

# Save the figure

agg_png("figures/results.png",
        width = 30, height = 25,
        units = "cm",
        res = 600,
        scaling = 1.8)
results_cor
dev.off()


# Results by species

species_results_cor <- ggscatter(flipper_mass_data, x = 'flipper_length_mm', y = 'body_mass_g', size = 1, color = 'species', palette = c(Adelie = "darkorange", Chinstrap = "purple", Gentoo = "cyan4"),
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "pearson",
          cor.coef.size = 2.5,
          xlab = "Flipper Length (mm)", ylab = "Body Mass (g)",
          facet.by = 'species', scales = "free_x")

species_results_cor

# Save the figure

agg_png("figures/speciesresults.png",
        width = 30, height = 25,
        units = "cm",
        res = 600,
        scaling = 1.8)
species_results_cor
dev.off()
```

### Conclusion

There is a positive correlation between penguin flipper length and body mass. Gentoo penguins are larger than Adelie or Chinstrap both in terms of body mass and flipper length. 
